---
title: "'R'mut Group Project"
author: ""
output: 
    html_document:
      code_folding: hide
      theme: united
---




## Key takeaways


We analyzed effect of TCMB interest rates on economy and our results :

*Interest rates don't work alone even if it's a big instrument to balance the dollar


## Overview and Preliminary

We get TCMB interest rate as a main data and USD/TRY.................... data it was a small data but very open the interpretation



Objectives :

 * Get raw data and arrange according to the result we want to learn
 * After put in order data visualize it and make INFERENCES
 * Find the areas most affected by interest rates
 
 
 
## Explore and Cleaning data

Our main dataset was small and clean but when we need comparisons other data we needed the make it compatible with main data.

First we need necessary libraries

```{r message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE)
library(readr)
library(dplyr)
library(lubridate)
library(ggplot2)
library(readxl)
```


```{r}
# Create a temporary file
tmp <- tempfile(fileext=".xlsx")

# Download file from repository to the temp file
download.file("https://github.com/pjournal/mef03g-r-mut/raw/master/project_datasets/tcmb_faiz.xlsx",destfile = tmp)
tcmb_int_table <- readxl::read_excel(tmp)

```

```{r message=FALSE, warning=FALSE}
tmp2 <- tempfile(fileext = ".csv")
download.file("https://github.com/pjournal/mef03g-r-mut/raw/master/project_datasets/USD_TRY.csv",destfile = tmp2)
usd_tl <- read_csv(tmp2)
```



```{r}
str(tcmb_int_table)
```

```{r}
usd_tl
```

```{r}
str(usd_tl)
```

I need convert date column format for easy using

```{r}
usd_tl_dp <- usd_tl %>% select("Date","Price") 

usd_tl_dp
```


```{r}
#I need convert date column format for easy using

usd_tl_dp$Date <- parse_date_time(usd_tl_dp$Date,orders = c("my"))

str(usd_tl_dp)

```



```{r}
graph1 <- ggplot(tcmb_int_table,aes(tcmb_int_table$year,tcmb_int_table$rate))+
geom_line(color="blue")+scale_y_continuous(
  "Interest Rate(%)",
  sec.axis = sec_axis(~ . * 0.33,name = "USD/TRY")
)+
  ggtitle("TCMB Interest Rates")+xlab("Years") +theme(plot.title = element_text(color = "cadetblue4",size=20))

graph1 <- graph1 +  geom_line(data=usd_tl_dp,aes(Date,Price*3),color="red") + ggtitle(("TCMB Interest Rates vs USD/TRY"))
graph1

```




```{r message=FALSE, warning=FALSE}
tmp3 <- tempfile(fileext = ".csv")
download.file("https://github.com/pjournal/mef03g-r-mut/blob/master/project_datasets/tech.csv",destfile = tmp3)
tech <- read_csv(tmp3)
```

```{r}
tech <- read.csv("tech.csv", header=TRUE, sep=",",skip = 1)
colnames(tech) <- c("country_name","country","date","indicator_name","indicator_code","value")

```


## Dot Plot the techonology data and see what it has
```{r}
library(scales)
theme_set(theme_classic())

# Plot
 ggplot(tech, aes(x=indicator_name, y=value)) + 
  geom_point(col='#e67e22', size=3) +   # Draw points
  geom_segment(aes(x=indicator_name, 
                   xend=indicator_name, 
                   y=min(value), 
                   yend=max(value)), 
               linetype="dashed", 
               size=0.1) +  # Draw dashed lines
  labs(title="Dot Plot", 
       subtitle="İndicators vs Values") +  
  coord_flip()
```
## Analysis of technology indicators
```{r}
ht <- tech%>%group_by(indicator_name,value,date)%>%filter(indicator_name == "High-technology exports (current US$)")%>%summarise(mean_value = mean(value/1000000))%>%arrange(desc(value))%>%group_by(date)
ch <- tech %>%group_by(indicator_name,value,date)%>%filter(indicator_name == "Charges for the use of intellectual property, payments (BoP, current US$)")%>%arrange(desc(value))%>%group_by(date)
```
```{r}
unht <- tech%>%group_by(date)%>% summarise(mean_value = mean(value/1000000))%>%arrange(desc(mean_value))%>%filter(date > 2010 & date < 2018)
```
##ompare between High tech and the other indicators
```{r}
all <- inner_join(ht,unht,by = c("date"))

ggplot(all, aes(x = date)) + 
  geom_line(aes(y = mean_value.x, colour = "High-technology exports")) +
  geom_line(aes(y = mean_value.y, colour = "other mean")) +
  xlab("Years") + ylab("Normalized Prices")  
```

```{r}
irate <- tcmb_int_table%>%mutate(date = year(year))
indicators <- tech %>% group_by(date,indicator_name)%>%filter(indicator_name %in% c("High-technology exports (current US$)", "Charges for the use of intellectual property, payments (BoP, current US$)","Charges for the use of intellectual property, receipts (BoP, current US$)")) %>%group_by(indicator_name)


all2 <- inner_join(irate,all,by = c("date"))
```

```{r}
g<-  ggplot(indicators, aes(x = date, y = value/100000000, colour=indicator_name)) + 
  geom_line() + 
  geom_point()+ 
  labs( x="" , y = "Tech Invest Value")+ 
  theme(legend.position = "bottom", legend.title = element_blank(), axis.text.x = element_text(size = 10)) 
g <- g +  geom_smooth(data=all2,aes(date,rate)) + ggtitle(("Tech Invest Rates vs Interest rates"))+  theme(axis.text.x = element_text(colour = "grey20", size = 10, angle = 45, hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(colour = "grey20", size = 10),
        text = element_text(size = 7),
        legend.position = "top", legend.box = "vertical" )
g

```

```{r}
theme_set(theme_bw())
g <- ggplot(all2, aes(date, value/100000000)) + 
  labs(title="TECH INVEST VS INTEREST RATE")

g + geom_jitter(aes(col=indicator_name, size=rate)) + 
  geom_smooth(aes(col=indicator_name), method="lm", se=F)+theme_light() +
  theme(axis.text.x = element_text(colour = "grey20", size = 10, angle = 45, hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(colour = "grey20", size = 10),
        text = element_text(size = 8),
        legend.position = "right", legend.box = "vertical" )
      
```



```{r}
graph1 <- ggplot(unht,aes(date,mean_value/35)) +ggtitle("Tech Invest Rates")+xlab("Years")+ylab("Rates(%)") +geom_smooth(se = FALSE, stat='summary',fun.y=quantile,fun.args=list(probs=0.9),color='blue')+theme(plot.title = element_text(color = "grey20",size=20))+scale_y_continuous(
  "Interest Rate(%)",
  sec.axis = sec_axis(~ . * 0.5,name = "Interest rates")
)
graph1 <- graph1 +  geom_smooth(se = FALSE, data=all2,aes(date,rate),color='red') + ggtitle(("Tech Invest Rates vs Interest rates"))+theme(plot.title = element_text(color = "grey20",size=20)) 
graph1
```




