---
title: "'R'mut Group Project"
author: ""
output: 
    html_document:
      code_folding: hide
      theme: united
---

## Key takeaways

We analyzed effect of TCMB interest rates on economy and our results :

*Interest rates don't work alone even if it's a big instrument to balance the dollar

## Overview and Preliminary

We get TCMB interest rate as a main data and USD/TRY.................... data it was a small data but very open the interpretation

## Objectives :
 
 * Get raw data and arrange according to the result we want to learn
 * After put in order data visualize it and make INFERENCES
 * Find the areas most affected by interest rates
 
## Explore and Cleaning data

Our main dataset was small and clean but when we need comparisons other data we needed the make it compatible with main data.
Firstly, we need necessary libraries:

```{r message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE)
library(readr)
library(dplyr)
library(lubridate)
library(ggplot2)
library(readxl)
library(tidyr)
library(plotly)
```
## Data Loading
```{r message=FALSE, warning=FALSE, include=FALSE}
# Create a temporary files to download our data
tmp <- tempfile(fileext=".xlsx")
tmp1 <- tempfile(fileext=".csv")
tmp2 <- tempfile(fileext=".xlsx")
tmp3 <- tempfile(fileext=".xlsx")
tmp4 <- tempfile(fileext=".csv")
tmp5 <- tempfile(fileext=".xlsx")
# Download file from repository to the temp file
download.file("https://github.com/pjournal/mef03g-r-mut/raw/master/project_datasets/tcmb_faiz.xlsx", destfile = tmp)
download.file("https://github.com/pjournal/mef03g-r-mut/raw/master/project_datasets/USD_TRY.csv", destfile = tmp1)
download.file("https://github.com/pjournal/mef03g-r-mut/raw/master/project_datasets/CPI.xlsx", destfile = tmp2)
download.file("https://github.com/pjournal/mef03g-r-mut/raw/master/project_datasets/PPI.xlsx", destfile = tmp3)
download.file("https://github.com/pjournal/mef03g-r-mut/raw/master/project_datasets/tech.csv",destfile = tmp4)
download.file("https://github.com/pjournal/mef03g-r-mut/raw/master/project_datasets/interest_rate_and_foreign_trade.xlsx",destfile = tmp5)
interest_rate <- readxl::read_excel(tmp)
usd_tl <- read_csv(tmp1)
consumer_price_index <- readxl::read_excel(tmp2)
producer_price_index <- readxl::read_excel(tmp3)
tech <- read.csv(tmp4, header = TRUE, sep = ",", skip = 1)
foreign_trade <- readxl::read_excel(tmp5)
```
Explore the interest_rate data frame with str() function.
```{r}
str(interest_rate) # to see dataframe format
```
```{r warning=FALSE}
str(usd_tl)
```
```{r}
str(consumer_price_index[1:5])
```
```{r}
str(producer_price_index[1:5])
```
## Data Type Converting
Since the "year" column appears a categorical variable on consumer price index, producer price index and interest rate datasets, it should be converted it to date with as.Date functon to more convenient analysis. Also, some columns are renamed to easy understanding.

```{r}
interest_rate <- interest_rate %>%
  mutate(date = year(year))
usd_tl_dp <- usd_tl %>% select("Date","Price") 
usd_tl_dp$Date <- parse_date_time(usd_tl_dp$Date,orders = c("my")) 
usd_tl_dp$Date <- as.Date(as.POSIXct(usd_tl_dp$Date))
consumer_price_index$Date <- as.Date(paste(consumer_price_index$Date,"-01",sep=""))
producer_price_index$Date <- as.Date(paste(producer_price_index$Date,"-01",sep=""))
colnames(tech) <- c("country_name","country","date","indicator_name","indicator_code","value")
```

A new dataset was prepared for changes in the CPI and PPI overall Turkey. At the same time, the 
dataset was manipulated by normalizing the datatable with gather() function.

```{r}
general_price_index <- data.frame(consumer_price_index$Date, consumer_price_index$Turkey, producer_price_index$Turkey)
names(general_price_index) <- c("Date", "CPI" , "PPI")
general_price_index <- general_price_index %>%
  gather(key = "index_type", "index_value", "CPI", "PPI")
general_price_index$index_type <- as.factor(general_price_index$index_type)
str(general_price_index)
```
## Analysis

Let's look at the graph of the interest rates change over the years.

```{r pressure, echo=TRUE, warning=FALSE, paged.print=TRUE}
c <- ggplot(interest_rate, aes(x=year, y=rate, group=1)) +
  geom_line(color = "blue") +
  geom_point(color = "blue") +
  labs(title = "TCMB Interest Rates",
       x = "Year of observation",
       y = "Interest Rates") + 
  theme_light() +
  theme(axis.text.x = element_text(colour = "grey0", size = 10, angle = 30, hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(colour = "grey20", size = 10),
        text = element_text(size = 12)) 
c <- ggplotly(c)
c
```

As can be seen from the graph, interest rates determined by TCMB showed volatility and increased momentum.

## USD/TRY relation
```{r}
interest_rate$year <- as.POSIXct(interest_rate$year)
usd_tl_dp$Date <- as.POSIXct(usd_tl_dp$Date)
graph_usdTL <- ggplot(interest_rate,aes(interest_rate$year,interest_rate$rate)) +
  geom_line(color="blue") +
  scale_y_continuous("Interest Rate(%)",
  sec.axis = sec_axis(~ . * 0.33, name = "USD/TRY")) +
  theme_light() +
  ggtitle("TCMB Interest Rates")+
  xlab("Years") +
  theme(plot.title = element_text(color = "cadetblue4", size=20))
graph_usdTL <- graph_usdTL +
  geom_line(data=usd_tl_dp, aes(Date,Price*3), color="red") + 
  ggtitle(("TCMB Interest Rates vs USD/TRY"))
graph_usdTL
```
## Inflation vs. Interest Rates

There are many discussions on the relationship between interest and inflation. Inflation refers to the rate at which prices for goods and services rise. To examine this relationship, let us examine the producer price index and consumer price index data. 


Although the graph values show an increasing change, it is also necessary to calculate the monthly and annual changes in interest rates in order to more convenient analysis.

```{r, echo=TRUE, message=TRUE, warning=FALSE, paged.print=TRUE}
general_price_index <- general_price_index %>% group_by(index_type) %>%
  mutate(monthly_change = (index_value / lag(index_value) -1)*100, yearly_change = (index_value / lag(index_value, 12) -1)*100) %>%
  gather(key = "change_type", "change_value", "monthly_change", "yearly_change")
```

We have changed the column names for easier visualizing.

```{r, echo=TRUE, message=TRUE, warning=FALSE, paged.print=TRUE}
general_price_index$change_type[general_price_index$change_type == "monthly_change"] <- "Monthly"
general_price_index$change_type[general_price_index$change_type == "yearly_change"] <- "Yearly"
```

The relationship between CPI and PPI is shown below.

```{r Graph, echo=TRUE, message=TRUE, warning=FALSE}
graph_index <- general_price_index %>% ggplot() +              # basic graphical object
  geom_line(aes(Date, index_value, color = index_type)) +
  labs(title = "Price Indexes",
       x = "Year of observation",
       y = "Indexes")+
  theme_light() +
  theme(axis.text.x = element_text(colour = "grey10", size = 10, angle = 45, hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(colour = "grey10", size = 10),
        text = element_text(size = 12),
        legend.title = element_text("Index Type"))
graph_index <- ggplotly(graph_index)
graph_index
```
Changes on PPI and CPI: 
```{r echo=TRUE, message=TRUE, warning=FALSE}
graph_index_change <- general_price_index %>% ggplot() +              
  geom_line(aes(x=Date, y=change_value, color = change_type), linetype = 0) +
  facet_wrap(general_price_index$index_type, dir = "v") +
  coord_cartesian()+
  labs(title = "Change in price indexes",
       x = "Year of observation",
       y = "Change")+
  theme_light() +
  theme(legend.title = element_text("Change Type"),
        axis.text.x = element_text(colour = "grey20", size = 10, hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(colour = "grey20", size = 10),
        text = element_text(size = 12))
graph_index_change <- ggplotly(graph_index_change)
graph_index_change
```
Our studies show that the inflation rate rises gradually over the years.

```{r, echo=TRUE, message=TRUE, warning=FALSE, paged.print=TRUE}
average <- general_price_index %>% 
  mutate(Year = year(Date)) %>%
  group_by(Year, index_type) %>% 
  summarize(average = mean(index_value))
graph_average <- ggplot(data = average) +
  geom_bar(mapping = aes(x=Year, y=average, fill = index_type), stat = "identity") +
  facet_wrap(average$index_type) +
  coord_cartesian()+
  labs(title = "Change in price indexes",
       x = "Year of observation",
       y = "Change") +
  theme_light() +
  theme(legend.title = element_text("Change Type"),
        axis.text.x = element_text(colour = "grey20", size = 10, hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(colour = "grey20", size = 10),
        text = element_text(size = 12))
graph_average <- ggplotly(graph_average)
graph_average
```
## Technologh Investment vs. Interest Rate
### Dot Plot the techonology data and see what it has

```{r warning=FALSE}
library(scales)
theme_set(theme_classic())
# Plot
ggplot(tech, aes(x=indicator_name, y=value)) + 
  geom_point(col='#e67e22', size=3) +   # Draw points
  geom_segment(aes(x=indicator_name, 
                   xend=indicator_name, 
                   y=min(value), 
                   yend=max(value)), 
               linetype="dashed", 
               size=0.1) +  # Draw dashed lines
  labs(title="Dot Plot",
       subtitle="İndicators vs Values") +  
  coord_flip()
```
### Analysis of technology indicators

```{r}
ht <- tech %>% 
  group_by(indicator_name, value, date) %>%
  filter(indicator_name == "High-technology exports (current US$)") %>%
  summarise(mean_value = mean(value/1000000)) %>% 
  arrange(desc(value)) %>% group_by(date)
ch <- tech %>% group_by(indicator_name,value,date) %>% 
  filter(indicator_name == "Charges for the use of intellectual property, payments (BoP, current US$)")%>%
  arrange(desc(value))%>%group_by(date)
```

```{r}
unht <- tech %>% group_by(date) %>%
  summarise(mean_value = mean(value/1000000)) %>% 
  arrange(desc(mean_value))%>%
  filter(date > 2010 & date < 2018)
```

### Compare between High tech and the other indicators

```{r}
all <- inner_join(ht,unht,by = c("date"))
ggplot(all, aes(x = date)) + 
  geom_line(aes(y = mean_value.x, colour = "High-technology exports")) +
  geom_line(aes(y = mean_value.y, colour = "other mean")) +
  xlab("Years") + ylab("Normalized Prices")  +
  theme_light()
```
```{r}
indicators <- tech %>% group_by(date, indicator_name) %>% 
  filter(indicator_name %in% c("High-technology exports (current US$)", "Charges for the use of intellectual property, payments (BoP, current US$)","Charges for the use of intellectual property, receipts (BoP, current US$)")) %>% 
  group_by(indicator_name)
all2 <- inner_join(interest_rate, all, by = c("date"))
```

```{r}
g <- ggplot(indicators, aes(x = date, y = value/100000000, colour=indicator_name)) + 
  geom_line() + 
  geom_point()+ 
  labs(x="",
       y = "Tech Invest Value") + 
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        axis.text.x = element_text(size = 10)) 
g <- g +  geom_smooth(data=all2, aes(date,rate)) + 
  ggtitle(("Tech Invest Rates vs Interest rates")) +
  theme(axis.text.x = element_text(colour = "grey20", size = 10, angle = 45, hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(colour = "grey20", size = 10),
        text = element_text(size = 7),
        legend.position = "top", legend.box = "vertical" )
g
```
```{r}
theme_set(theme_bw())
g <- ggplot(all2, aes(date, value/100000000)) + 
  labs(title="TECH INVEST VS INTEREST RATE")
g + geom_jitter(aes(col=indicator_name, size=rate)) + 
  geom_smooth(aes(col=indicator_name), method="lm", se=F)+theme_light() +
  theme(axis.text.x = element_text(colour = "grey20", size = 10, angle = 45, hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(colour = "grey20", size = 10),
        text = element_text(size = 8),
        legend.position = "right", legend.box = "vertical" )
      
```

```{r}
graph1 <- ggplot(unht,aes(date,mean_value/35)) +
  ggtitle("Tech Invest Rates")+xlab("Years")+ylab("Rates(%)") +
  geom_smooth(se = FALSE, stat='summary',
              fun.y=quantile,fun.args=list(probs=0.9),color='blue') +
  theme(plot.title = element_text(color = "grey20",size=20)) +
  scale_y_continuous("Interest Rate(%)",
                     sec.axis = sec_axis(~ . * 0.5,name = "Interest rates"))
graph1 <- graph1 +
  geom_smooth(se = FALSE, data=all2,aes(date,rate),color='red') + 
  ggtitle(("Tech Invest Rates vs Interest rates")) +
  theme(plot.title = element_text(color = "grey20",size=20)) 
graph1
```
## Relationship Between Interest Rate And Foreign Trade
The last day of each month were added as a date, the foreign trade figure of each month were added as a value and the valid interest rate for each date were added to rows for each date. Interest rates which did not coincide with any date were excluded from the dataset. 2010-05-31 is the first date and 2019-10-31	is the last date.

Assigning the columns in the dataset to variables to increase readability

```{r message = FALSE, warning = FALSE}
Exports <- foreign_trade$exports/1000000
Imports <- foreign_trade$imports/1000000
Balance <- foreign_trade$balance/1000000
Date <- foreign_trade$date
Rate <- foreign_trade$interest_rate
Deficit <- Balance*(-1)
```
### Graph of exports (USD in billions)

```{r message = FALSE, warning = FALSE}
ggplot(foreign_trade, aes(x = date, y = exports)) +
  geom_bar(stat ='identity', color = 'gray', fill = 'gray') +
  geom_line(stat ='identity', color = 'gray') +
  labs(title = 'Exports (USD in billions)', 
       caption="Source: TÜİK", 
       y = "Exports",
       x = "Year") +
  theme_classic()
```

### Graph of imports (USD in billions)

```{r message = FALSE, warning = FALSE}
ggplot(foreign_trade, aes(x = Date, y = Imports)) +
  geom_bar(stat ='identity', color = 'gray', fill = 'gray') +
  geom_line(stat ='identity', color = 'gray') +
  labs(title = 'Imports (USD in billions)', 
       caption="Source: TÜİK", 
       y = "Imports",
       x = "Year") +
  theme_classic()
```

### Graph of foreign trade balance (USD in billions)

```{r message = FALSE, warning = FALSE}
ggplot(foreign_trade, aes(x = Date, y = Balance)) +
  geom_bar(stat ='identity', color = 'brown', fill ='brown') +
  geom_line(stat ='identity', color = 'brown') +
  labs(title = 'Foreign Trade Balance (USD in billions)',
       caption="Source: TÜİK", 
       y = "Balance",
       x = "Year") +
  scale_y_reverse() +
  theme_classic()
```

### Comparison of interest rate and foreign trade deficit

```{r message = FALSE, warning = FALSE}
ggplot(foreign_trade, aes(x = Date)) +
  geom_line(aes(y = Deficit, color = 'CAD (USD in billions)')) +
  geom_line(aes(y = Rate, color = 'Interest Rate (%)')) +
  labs(title = 'Foreign Trade Deficit',
       caption="Source: TÜİK, TCMB", 
       y = "",
       x = "Year") +
  theme_classic()
ggplot(foreign_trade, aes(x = Date)) +
  geom_line(aes(y = Deficit, fill = 'CAD (USD in billions)'), color = 'red') +
  geom_smooth(aes(y = Deficit), 
              color = 'red', alpha = TRUE) +
  geom_line(aes(y = Rate), color = 'blue') +
  geom_smooth(aes(y = Rate, color = 'Rate', fill = 'Interest Rate (%)'), 
              color = 'blue', alpha = TRUE) +
  labs(title = 'Foreign Trade Deficit',
       caption="Source: TÜİK, TCMB", 
       y = "",
       x = "Year") +
  theme_classic()
```
